---
title: "Importing Data in R: Exploring Various Packages and Their Advantages "
author:
  - name:
      given: Ali
      family: Soleimani
    affiliation: Fresenius University of Applied Science
    url: https://alisoleimani71.github.io/
    email: soleimani.ali@stud.hs-fresenius.de
    corresponding: true
course: Data Analysis for Decision-Making
professor: Prof. Dr. Stephan Huber
date: today
date-format: long
format:
  html:
    embed-resources: true
    page-layout: article
    toc: true
a4paper: true
bibliography: "references.bib"
csl: "apa.csl"
reference-location: document
number-sections: true
---

<p style="text-align: center;">

Matriculation Nr.: 401166548

</p>

<p style="text-align: center;">

Email: [soleimani.ali@stud.hs-fresenius.de](mailto:Alisoleimani71@gmail.com){.email}

</p>

## why to import data?

Data import is the foundational step of any data analysis workflow. Before cleaning, transformation, modeling, or visualization can occur, data must be reliably loaded into R. As Wickham and Çetinkaya‑Rundel (2023) emphasize, *importing data is the gateway to all subsequent analysis*, and errors at this stage propagate throughout the entire pipeline.

The R ecosystem provides a rich set of tools for importing data from diverse formats, including CSV, Excel, JSON, XML, and statistical software files. The official tidyverse documentation highlights that choosing the correct import method improves reproducibility, performance, and data integrity (Posit, 2023).

## Why Data import is Important?

Data import is important for three academic reasons:

**1. Reproducibility**

Reliable import ensures that analyses can be replicated across systems, operating systems, and environments (Peng, 2011). Incorrect import leads to inconsistent results.

**2. Data Integrity**

Parsing errors—such as incorrect column types, encoding issues, or malformed rows—can introduce bias or invalidate results (Broman & Woo, 2018).

**3. Performance**

Efficient import tools reduce computation time, especially for large datasets. The fread() function from data.table is shown to outperform base R by an order of magnitude (data.table Project, 2024).

## different data types

the most common data types encountered in practice include:

1\. Text-based tabular data

-   CSV\
-   TSV\
-   Fixed-width files

2\. Spreadsheet data

-   Excel (.xls, .xlsx)

3\. Statistical software formats

-   SPSS (.sav)\
-   Stata (.dta)\
-   SAS (.sas7bdat)

4\. Semi-structured data

-   JSON\
-   XML

5\. Web-based data

-   APIs\
-   HTML tables\
-   Remote CSV files

6\. Databases

-   SQL-based systems (via DBI)

Each data type requires specialized import tools to preserve structure, metadata, and encoding.

## different packages for each data type

The mapping between data types and R packages is well-established in academic and official documentation:

|                   |                          |                             |
|-------------------|--------------------------|-----------------------------|
| **Package**       | **Best For**             | **Why**                     |
| readr             | CSV, TSV, delimited text | Fast, consistent parsing    |
| data.table::fread | Large CSV files          | Fastest import method       |
| readxl            | Excel (.xls, .xlsx)      | No dependencies, stable     |
| haven             | SPSS, Stata, SAS         | Preserves labels & metadata |
| jsonlite          | JSON, APIs               | Handles nested structures   |
| xml2              | XML                      | Robust parsing              |
| httr / rvest      | Web APIs, HTML scraping  | Flexible                    |

## Samples of Packages, With Codes for Each

### readr (CSV)

```{r}

library(readr)
df_readr <- read_csv("data.csv")
df_readr
```

### data.table::fread (Large CSV)

```{r}

library(data.table)
df_fread <- fread("bigdata.csv")
df_fread
```

### readxl (Excel)

```{r}
library(readxl)
df_excel <- read_excel("file.xlsx")
df_excel
```

## haven (SPSS/Stata/SAS)

```{r}
library(haven)
df_spss <- read_sav("survey.sav")
df_spss
```

### jsonlite (JSON)

```{r}
library(jsonlite)
df_json <- fromJSON("data.json")
df_json
```

### xml2 (XML)

```{r}
#| echo: false
library(xml2)
doc_xml <- read_xml("file.xml")
doc_xml
```

## importing data from multiple files

```{r}
#| eval: false
library(fs)
library(purrr)
library(readr)
library(dplyr)

files <- dir_ls("data/", glob = "*.csv")

df_multi <- files |>
  map(read_csv) |>
  bind_rows()

df_multi
```

## comparison table based on file type

| Format | Best Package | Strengths | Limitations |
|------------------|------------------|------------------|------------------|
| CSV | readr | Fast, tidyverse-friendly | Not ideal for huge files |
| Large CSV | fread | Fastest, auto-detects delimiter | Fewer diagnostics |
| Excel | readxl | No dependencies, reliable | Read-only |
| SPSS/Stata/SAS | haven | Preserves labels | SAS files slower |
| JSON | jsonlite | Handles nested data | Requires JSON knowledge |
| XML | xml2 | Robust parsing | Complex structures |
| Web APIs | httr | Flexible | Requires manual parsing |

## conclusion

Data import is not merely a technical step but a methodological foundation for reproducible and reliable analysis. R provides specialized tools for each data type, and selecting the correct package ensures optimal performance, data integrity, and workflow efficiency. By integrating the course notes, R4DS, and CRAN documentation, this handout provides a comprehensive overview of modern data import practices in R.

## Class assignment

Download the following dataset: https://raw.githubusercontent.com/allisonhorst/palmerpenguins/main/inst/extdata/penguins.csv

Tasks:

1\. Import the dataset using readr, fread, and base R.

2\. Measure import time using system.time().

3\. Compare column types produced by each method.

4\. Write a short paragraph explaining which method is best and why.

## References

-   **Broman, K., & Woo, K. (2018).** *Data organization in spreadsheets.* PLOS Computational Biology. DOI: https://doi.org/10.1371/journal.pcbi.1005746 [(doi.org in Bing)](https://www.bing.com/search?q=%22https%3A%2F%2Fdoi.org%2F10.1371%2Fjournal.pcbi.1005746%22)

-   **data.table Project. (2024).** *data.table: Fast data manipulation in R.* CRAN Documentation. (No DOI — official CRAN manual)

-   **Hubchev. (2024).** *Data Science Notes.* <https://hubchev.github.io/ds/>

-   **Ooms, J. (2014).** *The jsonlite package: A practical and consistent mapping between JSON data and R objects.* The R Journal. DOI: https://doi.org/10.32614/RJ-2014-011 [(doi.org in Bing)](https://www.bing.com/search?q=%22https%3A%2F%2Fdoi.org%2F10.32614%2FRJ-2014-011%22)

-   **Peng, R. (2011).** *Reproducible research in computational science.* Science. DOI: https://doi.org/10.1126/science.1213847 [(doi.org in Bing)](https://www.bing.com/search?q=%22https%3A%2F%2Fdoi.org%2F10.1126%2Fscience.1213847%22)

-   **Posit. (2023).** *Importing data with the tidyverse.* <https://r4ds.hadley.nz/data-import.html> (No DOI — official Posit documentation)

-   **R Project. (2024).** *Read data files from various sources.* CRAN Task View. (No DOI — official CRAN documentation)

-   **RStudio. (2024).** *readxl: Read Excel Files.* CRAN Documentation. (No DOI — official CRAN manual)

-   **Wickham, H. (2014).** *Tidy data.* Journal of Statistical Software. DOI: https://doi.org/10.18637/jss.v059.i10 [(doi.org in Bing)](https://www.bing.com/search?q=%22https%3A%2F%2Fdoi.org%2F10.18637%2Fjss.v059.i10%22)

-   **Wickham, H., & Çetinkaya‑Rundel, M. (2023).** *R for Data Science (2nd ed.).* Posit. <https://r4ds.hadley.nz> (No DOI — open textbook)

:::
